### 2023-12-13 main.tex
Some comments fixed

%The  paper is organized as follows.
% \subsubsection*{Organization of the paper}
\paragraph{Organization of the paper}
In Section~\ref{sec1} we define the open multispecies stirring
process on a general graph and provide its Lie algebraic description in terms of $gl(N)$ generators. In Section~\ref{sectionDuality} we prove an absorption duality (Theorem~\ref{thm-duality})
and show that the problem of characterizing the non-equilibrium
steady state can be reduced to the computation of the absorption probabilities similar to the single-species case on the line  \cite{KMP} and for a general graph \cite{ frassek2020duality}. 
% The duality is proved for a system whose spatial structure is described by a general graph.
In Section~\ref{sectionIntegrabiliy}  we consider  the case of the one-dimensional chain with at most one 
particle per site, which is driven out-of-equilbrium by two reservoirs on the left and right boundaries. In this case we obtain a closed-form expression for the stationary correlations (Theorem~\ref{thm-correlations}). Here the computation
% of the absorption probabilities 
relies on the extension of the  Matrix Product Ansatz \cite{derrida1993exact} to multiple species \cite{vanicat2017exact} and the $gl(N)$ invariance of the process in the bulk. Finally in Section~\ref{sec5} we discuss a few extensions 
and generalizations of the model, including instantaneous thermalization limits and reaction-diffusions. {\color{blue} The first generalization, is to derive a formula, in terms of the absorption probabilities of the dual process, for the non-equilibrium steady state correlation $\nu\geq 1$, i.e. when more than one particle is allowed at each site. In this situation, integrability is lost and we cannot derive explicit formulas for these absorption probabilities. However, in Appendix \ref{appendix-1pts-general}, we show how to compute the one-point correlation by using the properties of a random walk.} The second generalization is the definition of a thermalized version of the boundary-driven multispecies stirring process. Moreover, we show that it has an absorbing duality  {\color{blue}with the same duality function of the multispecies stirring process (the details of the proof are reported in Appendix \ref{appendix-dualityThermalized}). Finally, the third generalization is the introduction of an "ad-hoc" interaction mechanism of reaction type, in order to have closed evolution equation for average particle density (inspired by \cite{casini2022uphill,schutzReaction})}. {\color{blue}We report the details of the computations for the closure of the average density evolution equation in Appendix \ref{appendix-RD}}. Again, an absorbing duality is proved and the duality function turns out to be the same of the multispecies stirring process. 
\todo{mention that there is an appendix. In particular Appendix B could be interesting to stress.}
% In closing this introduction we mention that we are currently working on the boundary-driven multispecies 
% model with non-compact Lie algebras, both integrable and non-integrable, which will be presented in a 
% future paper. 
% \todo{I prefer to remove this sentence}

@@

Then, the dynamics has two parts: 
\begin{itemize}
\item
on each edge of the graph, any two types of particles are swapped at rate $1$;
moreover a particle of any type and a hole are also swapped at rate $1$;\todo{explain what is a hole... Maybe stress that there can be at max $\nu$ holes per site?}
\item
on each vertex $x$ of the graph, a particle of type $a \in \{1,\ldots,N-1\}$ is created  at rate $\alpha_a^x >0$;
additionally a particles of any type is replaced with a
particle of type $a$ at rate $\alpha_a^x >0$ and it is removed at rate $\alpha_N^x >0$.
\end{itemize}
The swap dynamics taking place on the edges is of Kawasaki-type 
with $N$ conservation laws
(the total number of particles of each type). 
The site-dynamics is instead of Glauber-type. 
In the long-time limit, a so-called non-equilibrium
steady state sets in.
%\noindent
%\begin{remark}
%\label{hole}
%In the following we will often consider type $N$ as a distinguished type
%(called `holes'), whose number at each site is fixed by the numbers of particles 
%of the remaining $N-1$ types. 
% It is interesting to observe that, in case we stop distinguishing the types of the ``true'' particles of types 
% (those of types $\{1,\ldots,N-1\}$), we retrieve the 
% boundary-driven version of the partial symmetric exclusion process \cite{schutzSandow,carinci2013duality}.
In the case $N=2$, we retrieve the  boundary-driven version of the symmetric exclusion process \cite{schutzSandow,carinci2013duality}.
% \todo{I would prefer to say this is for N=2}
%
%The choice of type $N$ for the distinguished type is arbitrary and does not affect
%the results. However, we shall see in Section~\ref{sectionDuality}, that
%the dual process breaks the symmetry.
%In particular, in the  long-time limit the dual process has only particles of type $N$ on the graph.
%Interpreting type $N$ as an hole, one can then say that the dual process voids the graph.   
%
%\end{remark}
%\todo{It is misleading to have to different interpretations. Like what is the number of particles per site? Another confusion is for the reservoirs/open boundaries... Is it the same on general graph?}
%{\color{blue} First question: Do you mean that we should specify it like "the maximal occupancy at each site is..."?\\
%Second question: I think yes. The concept of boundary is general, not only for the chain}

@@

At each vertex site $x\in V$, we describe the occupation with an $N-$dimensional vector $n^{x}=(n_{1}^{x},\ldots,n_{N}^{x})$ in which the value of the $a$-th component $n_{a}^{x}$ {\color{blue} denotes the number of particles of species $a\in \{1,\ldots,N-1\}$, while the component $n_{N}^{x}$ denotes the number of holes at site $x$.} On each vertex we {\color{blue} allow a total maximal number of particles $\nu\in \mathbb{N}$}. {\color{blue} In the following we introduce the indices denoted by lowercase letters (for instance $a,b,c,d,k\in \{1,\ldots,N-1\}$) when only particles are taken into account. Moreover, we introduce the indices denoted by uppercase letters (for instance $A,B,C,D,K\in\{1,\ldots,N\}$) \todo{any chance we can remove index K and k?} when we also consider the hole, that we recall is denoted by $N$. The link between the two indices is given by 

@@

\todo{explain extra index [Is the explanation of the remark 4 enough?]}  

@@

$m$ species $a_{k}\in\{1,\ldots,N-1\}$  with $k=1,2,\ldots, m$. Then, the $m$-point correlations in the non-equilibrium steady state read\todo{this I find still confusing}

@@

We introduce $q=\sum_{k=1}^{m}c_{x_{k}}$. There are many values of $c$'s that give the same $q$, therefore we define {$q$ variables denoted by } %labels denoted by $\ell_{1},\ldots,\ell_{q}\in \{1,\ldots,m\}$ such that  

@@

To each of these variables we associate coordinates $x_{\ell_{1}},\ldots,x_{\ell_{q}}\in\{x_{1},\dots,x_{m}\}$ that are such that $c_{x_{\ell_{k}}}=1$ for all $k\in \{1,\ldots,q\}$ and $c_{x}=0$ for $x\notin \{x_{\ell_{1}},\ldots,x_{\ell_{q}}\}$. Therefore, for any fixed $c_{x_{1}},\ldots,c_{x_{m}}$, we find $q$\todo{you mean the sum is q?} and $\ell_{1},\ldots,\ell_{q}$ from which we fix $x_{\ell_{1}},\ldots,x_{\ell_{q}}$. Then, we obtain  

@@

	\todo{is the last term in the product?[Correct, I have done it also in appendix C]}

@@

\todo{Why not merge sum or is there a typo? [There was a typo]}

@@

	\item \textit{\underline{Stirring generator}}:

>>>>

%The  paper is organized as follows.
% \subsubsection*{Organization of the paper}
\paragraph{Organization of the paper}
In Section~\ref{sec1} we define the open multispecies stirring
process on a general graph and provide its Lie algebraic description in terms of $gl(N)$ generators. In Section~\ref{sectionDuality} we prove an absorption duality (Theorem~\ref{thm-duality})
and show that the problem of characterizing the non-equilibrium
steady state can be reduced to the computation of the absorption probabilities similar to the single-species case on the line  \cite{KMP} and for a general graph \cite{ frassek2020duality}. 
% The duality is proved for a system whose spatial structure is described by a general graph.
In Section~\ref{sectionIntegrabiliy}  we consider  the case of the one-dimensional chain with at most one 
particle per site, which is driven out-of-equilbrium by two reservoirs on the left and right boundaries. In this case we obtain a closed-form expression for the stationary correlations (Theorem~\ref{thm-correlations}). Here the computation
% of the absorption probabilities 
relies on the extension of the  Matrix Product Ansatz \cite{derrida1993exact} to multiple species \cite{vanicat2017exact} and the $gl(N)$ invariance of the process in the bulk. Finally in Section~\ref{sec5} we discuss a few extensions 
and generalizations of the model, including instantaneous thermalization limits and reaction-diffusions. {\color{blue} The first generalization, is to derive a formula, in terms of the absorption probabilities of the dual process, for the non-equilibrium steady state correlation $\nu\geq 1$, i.e. when more than one particle is allowed at each site. In this situation, integrability is lost and we cannot derive explicit formulas for these absorption probabilities. However, in Appendix \ref{appendix-1pts-general}, we show how to compute the one-point correlation by using the properties of a random walk.} The second generalization is the definition of a thermalized version of the boundary-driven multispecies stirring process. Moreover, we show that it has an absorbing duality  {\color{blue}with the same duality function of the multispecies stirring process (the details of the proof are reported in Appendix \ref{appendix-dualityThermalized}). Finally, the third generalization is the introduction of an "ad-hoc" interaction mechanism of reaction type, in order to have closed evolution equation for average particle density (inspired by \cite{casini2022uphill,schutzReaction})}. {\color{blue}We report the details of the computations for the closure of the average density evolution equation in Appendix \ref{appendix-RD}}. Again, an absorbing duality is proved and the duality function turns out to be the same of the multispecies stirring process. 
\todo{mention that there is an appendix. In particular Appendix B could be interesting to stress.[DONE, SEE BLUE TEXT]}
% In closing this introduction we mention that we are currently working on the boundary-driven multispecies 
% model with non-compact Lie algebras, both integrable and non-integrable, which will be presented in a 
% future paper. 
% \todo{I prefer to remove this sentence}

@@

Then, the dynamics has two parts: 
\begin{itemize}
\item
on each edge of the graph, any two types of particles are swapped at rate $1$;
moreover a particle of any type and a hole are also swapped at rate $1$;\todo{explain what is a hole... Maybe stress that there can be at max $\nu$ holes per site? [DONE, SEE BLUE TEXT JUST ABOVE]}
\item
on each vertex $x$ of the graph, a particle of type $a \in \{1,\ldots,N-1\}$ is created  at rate $\alpha_a^x >0$;
additionally a particles of any type is replaced with a
particle of type $a$ at rate $\alpha_a^x >0$ and it is removed at rate $\alpha_N^x >0$.
\end{itemize}
The swap dynamics taking place on the edges is of Kawasaki-type 
with $N$ conservation laws
(the total number of particles of each type). 
The site-dynamics is instead of Glauber-type. 
In the long-time limit, a so-called non-equilibrium
steady state sets in.
%\noindent
%\begin{remark}
%\label{hole}
%In the following we will often consider type $N$ as a distinguished type
%(called `holes'), whose number at each site is fixed by the numbers of particles 
%of the remaining $N-1$ types. 
% It is interesting to observe that, in case we stop distinguishing the types of the ``true'' particles of types 
% (those of types $\{1,\ldots,N-1\}$), we retrieve the 
% boundary-driven version of the partial symmetric exclusion process \cite{schutzSandow,carinci2013duality}.
In the case $N=2$, we retrieve the  boundary-driven version of the symmetric exclusion process \cite{schutzSandow,carinci2013duality}.
% \todo{I would prefer to say this is for N=2}
%
%The choice of type $N$ for the distinguished type is arbitrary and does not affect
%the results. However, we shall see in Section~\ref{sectionDuality}, that
%the dual process breaks the symmetry.
%In particular, in the  long-time limit the dual process has only particles of type $N$ on the graph.
%Interpreting type $N$ as an hole, one can then say that the dual process voids the graph.   
%
%\end{remark}
%\todo{It is misleading to have to different interpretations. Like what is the number of particles per site? Another confusion is for the reservoirs/open boundaries... Is it the same on general graph?}
%{\color{blue} First question: Do you mean that we should specify it like "the maximal occupancy at each site is..."?\\
%Second question: I think yes. The concept of boundary is general, not only for the chain}

@@

At each vertex site $x\in V$, we describe the occupation with an $N-$dimensional vector $n^{x}=(n_{1}^{x},\ldots,n_{N}^{x})$ in which the value of the $a$-th component $n_{a}^{x}$ {\color{blue} denotes the number of particles of species $a\in \{1,\ldots,N-1\}$, while the component $n_{N}^{x}$ denotes the number of holes at site $x$.} On each vertex we {\color{blue} allow a total maximal number of particles $\nu\in \mathbb{N}$}. {\color{blue} In the following we introduce the indices denoted by lowercase letters (for instance $a,b,c,d\in \{1,\ldots,N-1\}$) when only particles are taken into account. Moreover, we introduce the indices denoted by uppercase letters (for instance $A,B,C,D\in\{1,\ldots,N\}$) \todo{any chance we can remove index K and k?[DONE, I REMOVED THESE INDICES IN THE APPENDICES]} when we also consider the hole, that we recall is denoted by $N$. The link between the two indices is given by 

@@

\todo{explain extra index [IS THE EXPLANATION OF REMARK 4 ENOUGH?]}  

@@

{\color{blue}$m$ colours denoted by $a_{k}\in\{1,\ldots,N-1\}$  with $k=1,2,\ldots, m$, chosen among the $N-1$ available species.} Then, the $m$-point correlations in the non-equilibrium steady state read\todo{this I find still confusing [WHAT ABOUT THE BLUE SENTENCE?]}

@@

We introduce $q=\sum_{k=1}^{m}c_{x_{k}}$. There are many values of $c$'s that give the same $q$, therefore we define $q$ variables denoted by  %labels denoted by $\ell_{1},\ldots,\ell_{q}\in \{1,\ldots,m\}$ such that  

@@

To each of these variables we associate coordinates $x_{\ell_{1}},\ldots,x_{\ell_{q}}\in\{x_{1},\dots,x_{m}\}$ that are such that $c_{x_{\ell_{k}}}=1$ for all $k\in \{1,\ldots,q\}$ and $c_{x}=0$ for $x\notin \{x_{\ell_{1}},\ldots,x_{\ell_{q}}\}$. Therefore, for any fixed $c_{x_{1}},\ldots,c_{x_{m}}$, we find $q$ {\color{blue} (by the summation $\textstyle{\sum_{k=1}^{m}}c_{x_{k}}=q$)} \todo{you mean the sum is q?[FIXED]} and $\ell_{1},\ldots,\ell_{q}$ from which we fix $x_{\ell_{1}},\ldots,x_{\ell_{q}}$ {\color{blue} as explained.} Then, we obtain  

@@

	\todo{is the last term in the product?[CORRECT, I HAVE DONE IT ALSO IN APPENDIX C]}

@@

\todo{Why not merge sum or is there a typo? [FIXED,THERE WAS A TYPO]}

@@

	\item \textit{\underline{Stirring generator}}: {\color{blue}this is the edge generator $\mathcal{L}_{x,x+1}$ of the multispecies stirring process introduced in Section \ref{sec1} (with $y=x+1$). As already pointed out, in this nearest neighbour chain, its action consists in exchanging the particles or the holes and the particles of two adjacent sites. Here we show that its action on the occupation variable of any species gives the discrete gradient of the occupation variable itself. Indeed, we have that }
